The MACE is essentially unidimensional: Bifactor models of the `Maltreatment and Abuse Chronology of Exposure' scale

### 2022-08-06 manuscript/main.tex
getting close

@@

\abstract{Your abstract here.}

@@

Childhood maltreatment is an unfortunately common experience for children across the globe \citep{stoltenborgh2015prevalence}. It is a risk factor for multiple poor outcomes later in life, including the onset of mental illness \citep{macmillan2001childhood, green2010childhood, kessler2010childhood}, challenges in cognitive functioning \citep{kavanaugh2017neurocognitive, r2018common, su2019does}, and poorer physical health \citep{wegman2009meta, widom2012prospective, goodwin2004association}. Childhood maltreatment also predicts worse treatment outcomes for many psychiatric disorders \citep{nanni2012childhood, thomas2019childhood, schuckher2019history}. Thus, the reliable and valid assessment of childhood maltreatment is an important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design better interventions to prevent maltreatment in the first place.

There already exist many measures of childhood abuse and maltreatment for researchers to choose from \citep{saini2019systematic}. One recently developed measure, the Maltreatment and Abuse Chronology of Exposure (MACE) scale \citep{teicher2015maltreatment}, has a number of notable advantages. The MACE is a 52-item retrospective self-report questionnaire comprised of 10 short subscales that each assess a unique type of childhood maltreatment. Six of these (e.g. parental verbal abuse, physical abuse, nonverbal emotional abuse, sexual abuse, emotional neglect, physical neglect) overlap with other common measures of abuse; the remaining four assess forms of maltreatment that are less frequently represented in other measures (e.g. peer victimization, witnessing domestic violence). The items in the MACE were carefully selected to measure adverse childhood experiences of increasing severity in order to more finely differentiate people of varying exposure levels. The MACE also measures the timing of maltreatment experiences, which is critical for investigating how the onset and duration of abuse shapes trajectories of development. The MACE provides two measures, an overall severity score (number of maltreatment experiences) and a multiplicity score (number of types of maltreatment experienced), with excellent temporal stability, good convergent validity with other standard measures, and improved predictive validity in predicting psychiatric symptom ratings \citep{teicher2015maltreatment}. For these reasons, the MACE has been listed as among the best measures of childhood maltreatment currently available \citep{saini2019systematic, georgieva2022systematic}.

In recent years, there have been calls for childhood maltreatment research to move away from the cumulative risk framework \citep{evans2013cumulative} --- which focuses only on the number of adverse experiences --- towards a dimensional approach, which emphasizes how distinct types of maltreatment uniquely shape developmental trajectories \citep{mclaughlin2016beyond, belsky2012beyond}. In response, some researchers have begun to use the MACE to form new indices of maltreatment in lieu of or in addition to the severity and multiplicity scores. For example, some have investigated the unique contributions of individual differences on all 10 MACE subscale scores to the development of psychopathology \citep{schalinski2015type, gerke2018childhood, schalinski2019early}. Following a prominent dimensional model of maltreatment \citep{mclaughlin2014childhood}, other groups have made composite indices of threat and deprivation experiences using items in the MACE that respectively assess abuse and neglect experiences \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}. Though these alternatives maltreatment scores may yield new insight, their psychometric properties have not been extensively investigated.

One critical challenge in measuring childhood maltreatment is the high co-occurrence adverse experiences. That is, children experiencing one form of maltreatment (e.g. physical abuse) are likely to experience one or more other types (e.g. emotional abuse; \citealt{green2010childhood, mclaughlin2012childhood, marques2021risk, mersky2017rethinking}). As a consequence, isolating the unique contribution of one type of maltreatment to the development of some negative outcome is difficult insofar as any one type of adversity is likely confounded with general maltreatment. Concretely, this means that researchers who want to study the effects of particular types of maltreatment using subscale scores or composite indices need to verify those measures reflect a construct mostly independent of overall maltreatment, lest they draw spurious conclusions.

Bifactor models, one type of hierarchical factor models, are well-suited for addressing this issue \citep{bornovalova2020appropriate}. They assume that the covariance structure of a set of items can be accounted for by a general factor, reflecting shared variance across all items, and one or more specific factors, reflecting additional shared variance among subsets of items \citep{Reise2012-ql}. Crucially, bifactor models can be used to calculate model-based reliability statistics, including the omega hierarchical ($\omega_h$) and omega subscale ($\omega_s$) indices, which respectively measure the proportion of variance in total and subscale scores attributable to the general factor and specific factors \citep{reise2013applying, rodriguez2016evaluating}. Applied to the MACE scale, these indices can inform researchers as to whether subscale scores or composite indices for particular types of maltreatment are interpretable as such, or if they instead predominantly reflect overall maltreatment.

@@

The participants in the current study (N=1633 total) belonged to one of two samples. The first sample of participants (N=1051) are from the original development of the MACE scale \citep{teicher2015maltreatment}. The full recruitment details for this sample have been reported elsewhere \citep{teicher2015maltreatment}. Briefly, these participants were recruited from the greater Boston area between 2010 and 2013. Inclusion criteria included being medically healthy, unmedicated, and between 18–25 years of age.  This sample will hereafter be referred to as the \textit{original} sample.

@@

The demographics of the two samples are summarized in Table \ref{tab:demographics}. In both samples, the majority of participants identified as women; however, the original sample was comprised of proportionally more woman than the replication sample ($z$ = 4.142, $p$ < 0.001). Moreover, participants in the original sample were younger on average than participants in the replication sample ($t$ = -19.313, $p$ < 0.001). Finally, the majority of participants in both samples identified as White, and not as Hispanic or Latino. The proportion of participants identifying as White was not statistically different across samples ($z = 0.417$, $p = 0.677$); neither was the proportion of participants identifying as Hispanic or Latino ($z$ = -1.241, $p$ = 0.214).

@@

All participants completed the 52-item version of the MACE scale \citep{teicher2015maltreatment}. The scale is a retrospective measure of the severity and timing of exposure to 10 types of childhood maltreatment. For each item, participants endorse whether they ever experienced a particular event during childhood (e.g. ``Parents or guardians intentionally pushed, grabbed, shoved, slapped, pinched, punched or kicked you'') and, if so, at what ages (up to age 18). Here we only analyze the binary endorsement responses (Yes = 1, No = 0) and not the chronology data. Six of the 52-items --- evenly divided between the emotional and physical neglect subscales --- are positively valenced (e.g. ``One or more individuals in your family helped you feel important or special'') and reverse-scored. For these items we follow the scoring convention set by \cite{teicher2015maltreatment}: a score of 1 is assigned to a response only in the absence of a positive event for all 18 years of childhood. This is in contrast to the remaining 46 items in the scale, where a score of 1 is assigned to a response if maltreatment or abuse occurred in at least one year of childhood.

In both the original and replication samples, participants completed a number of additional self-report symptom measures. We do not include these in our analyses and therefore do not discuss these measures further in the main text. For completeness, we report these secondary self-report measures in the Supplementary Materials.

@@

To ensure data quality in the replication sample, we relied on attention checks embedded in the self-report measures to identify and remove participants engaging in careless or insufficient effort responding \citep{zorowitz2021inattentive}. The data from N=64 (9.9\%)  participants in the replication sample were excluded for failing one or more of these attention checks, leaving the data from N=582 participants for analysis.

@@

Prior to analysis, we investigated whether it was appropriate to combine response data from the original and replication samples. We first compared maltreatment endorsement rates for each item across samples. The Spearman-rank correlation of endorsement rates across samples was large ($\rho$ = 0.951, p < 0.001), indicating excellent correspondence in the relative prevalence of childhood maltreatment across samples. When we compare the distribution of person-level sum scores across samples,  however, we found that the number of maltreatment exposures was greater on average in the replication sample than in the original sample ($p_1$ = 0.185, $p_2$ = 0.219, $t$ = -4.611, p < 0.001).

A mean shift in overall endorsement rates is not necessarily an indication that measurement invariance has been violated. One possibility is simply that participants in the replication sample had simply experienced more childhood maltreatment than participants in the original sample. A second possibility is that the differences in the demographic composition of the two samples (e.g. gender, age) moderate the differences in endorsement rates. To investigate these possibilities, we tested for uniform differential item functioning (DIF) using logistic regression \citep{rogers1993comparison}, in which a participant's endorsement response for an item was predicted by their rest score (i.e. their sum score for all other items), sample membership, gender, and age.

Using the cutoffs recommended by \cite{hidalgo2014binary}, we identified 11 of 52 items on the MACE that exhibited large DIF by study (Figure \ref{fig:dif_study}). Interestingly these included all six of the reverse-scored items. We also found an additional 11 items that exhibited large DIF by gender (Figure \ref{fig:dif_gender}). Five of these items concerned sexual abuse (endorsed by more women than men), and another five concerned peer physical abuse (endorsed by more men than women). We identified no items that exhibited large DIF by age. Given the number of items exhibiting DIF by sample membership, we elected to model each sample separately.

@@

The MACE contains a total of 12 item pairs and triplets characterized by response dependence (Table \ref{tab:dependence}). The items in some of these sets are structurally dependent, where a response to one item requires, or strongly implies, a certain response on the second. For example, a person endorsing item 8 (``Parents or guardians hit you so hard that you received medical attention'') is all but certain to endorse item 7 (``Parents or guardians hit you so hard that it left marks for more than a few minutes''). Other items are dependent by virtue of redundancy. One such example is item 36 (``Peers forced you to engage in sexual activity against your will'') and 37 (``Peers forced you to do things sexually that you did not want to do``); here the same question is asked twice using slightly different language.

For this study, we were concerned about response dependence for two reasons. First, it can inflate the estimates of item discrimination parameters, making items appear more informative than they actually are \citep{marais2008formalizing}. Second, response dependence can inflate the estimates of item loadings on specific factors in bifactor models, giving false support for a larger number of factors than necessary \citep{reise2013applying}. Following \cite{marais2008formalizing}, we therefore converted each set of dependent items into a single polytomous item. Thus, the MACE was reduced from 52 to 39 items in total.

@@

To begin our investigation of the MACE factor structure, we fit a series of confirmatory item factor models to the response data from each sample separately. The basis of each model is the graded response model for ordered polytomous data \citep{samejima1997graded}. The factor structure for each of the four models is depicted in Figure \ref{fig:models}. The first model we fit is a one-factor (undimensional) model whereby all items load onto a general maltreatment factor (Figure \ref{fig:models}a). The second is a bifactor model, in which all items load onto a general maltreatment factor and also one of 10 specific factors corresponding to their subscale membership (Figure \ref{fig:models}b).

The third model we fit investigated a conceptual model of maltreatment that distinguishes between threat and deprivation as distinct forms of childhood adversity \citep{mclaughlin2014childhood}. Initially we fit a bifactor model in which all items loaded onto a general maltreatment factor one either the threat or deprivation specific factors. However, we observed vanishing item loadings on the threat specific factor. As such, we instead fit a bifactor S-1 model \citep{eid2017anomalous} where all items load onto a general maltreatment (or reference) factor, and only the items measuring neglect experiences load onto their own specific factor (Figure \ref{fig:models}c).

@@

The fourth and final model we fit was designed provide a more parsimonious factor structure for the MACE. Items in the MACE can be neatly divided according to the source of adversity: maltreatment by adults at home (e.g. parents) or by peers. We hypothesized then that there would be two principal factors underlying responses on the MACE, maltreatment by adults and peer victimization. Based on preliminary model fits, we also suspected the presence of a secondary methods factor affecting the 6 reverse-scored items. Initially we fit a bifactor model in which all items loaded onto a general maltreatment factor and also one of three specific factors: maltreatment by adults, peer victimization, or reverse-scoring. However, we also observed vanishing loadings for the maltreatment by adults specific factor. Thus, we instead fit a bifactor S-1 model where all items load onto a general maltreatment factor; in addition, the peer victimization and reverse-scoring items load onto their own respective specific factors (Figure \ref{fig:models}d).

@@

We relied on multiple indices to evaluate the fit of the models to the data. First we calculated three traditional goodness-of-fit statistics using the ordinal $M_2$ statistic \citep{cai2013limited, maydeu2013goodness}. $M_2$ is asymptotically $\chi^2$-distributed with $\kappa − \nu$ degrees of freedom ($df$), where $\kappa$ is the reduced number of first- and second-order marginal residuals (i.e. $\kappa = n(n + 1)/2$ for $n$ items) and $\nu$ denotes the number of free parameters. To improve the stability of the $M_2$ statistic, the responses to polytomous items were binarized (i.e. none vs. some maltreatment). Using the $M_2$ statistic, we calculated the root mean square error of approximation (RMSEA), comparative fit index (CFI), and Tucker-Lewis index (TLI) for each model fit. Following convention \citep{hu1999cutoff}, the benchmarks for adequate model fit are RMSEA < 0.08, CFI > 0.90, and TLI > 0.90; in turn, the benchmarks for good model fit are RMSEA < 0.05, CFI > 0.95, and TLI > 0.95.

Given the questionable utility of SEM-based fit indices for item response models (e.g. \citealt{reise2014evaluating}), we calculated two additional fit indices. First, we employed posterior predictive model checking and the $\chi^2_{NC}$ discrepancy measure based on the overall severity score distribution \citep{sinharay2006posterior}. This measure compares the observed and model-predicted proportion of participants at each overall severity score level. Second, we calculated Yen's $Q_3$ statistic for each pair of items to detect local dependence \citep{yen1984effects}. For each model and sample, we also calculated the critical $Q_3$-value above which local dependence is indicated  \citep{christensen2017critical}. Briefly, for each model and sample, we simulated 1000 locally independent datasets using the posterior distribution of the model parameters. We then recorded the maximum $Q_3$ value per simulation. The critical $Q_3$-value was defined as the 99th percentile of these max $Q_3$ values.

@@

To evaluate the reliability of the MACE total and subscale/composite scores, we computed several model-based reliability indices using the factor loading estimates from the bifactor models. First, we calculated coefficient omega ($\omega$; \citealt{mcdonald1999test}), which quantifies the proportion of reliable score variance attributable to all factors (i.e. both group and specific factors). Next we calculated coefficient omega hierarchical ($\omega_h$) and coefficient omega subscale ($\omega_s$;  \citealt{reise2013applying, rodriguez2016evaluating}). Whereas $\omega_h$ measures the fraction of total score variance attributable to the general factor, $\omega_s$ measures the fraction of subscale score variance attributable to a specific factor after controlling for the general factor. The values of $\omega_h$/$\omega_s$ facilitate interpretation of total and subscale scores. Values of $\omega_h$ approaching 1 indicate the general factor is the primary source of reliable variance in a total score. When $\omega_h$ > 0.80, the total score may be considered an essentially unidimensional reflection of the general factor (even if the data are multidimensional;  \citealt{rodriguez2016applying}). So too values of $\omega_s$ approaching 1 indicate a specific factor, and not the general factor, is the primary source of reliable variance in a subscale score. \cite{canivez2016bifactor} suggests that an acceptable $\omega_s$ value is 0.50 and that >0.70 is desirable. In contrast, values of $\omega_s$ smaller than 0.50 preclude meaningful interpretation of subscale scores as reflecting a specific factor \citep{gignac2013bifactor}.

To assess the unidimensionality of the MACE scale, we also calculated the explained common variance (ECV; \citealt{sijtsma2009use}) and the proportion of uncontaminated correlations (PUC; \citealt{reise2013multidimensionality}). The ECV is the ratio of the variance explained by the general factor divided by the total common variance of a scale; it therefore measures the contribution of the general factor relative to the specific factors. As ECV values approach 1, the general factor loadings of a bifactor model are expected to resemble those that would be obtained by a one-dimensional model. ECV values greater than 0.70 often indicate a scale is essentially unidimensional \citep{rodriguez2016applying}. In turn, the PUC quantifies how many inter-item correlations are accounted for only by the general factor and is an important moderator of the ECV. As the PUC increases, ECV is less important for evaluating the risk of bias when fitting a unidimensional model to data with a latent bifactor structure. When PUC > 0.80, there is a low risk of bias when a multidimensional scale is treated as unidimensional \citep{reise2013multidimensionality}. As an additional check, we calculated the relative parameter bias (RPB), which is the difference between an item's loading in a unidimensional model and its corresponding general factor loading the bifactor, divided by the general factor loading in the bifactor. Parameter bias less than 10–15\% is acceptable \citep{muthen1987structural}.

Finally, we calculated the H index to assess the construct replicability of each factor \citep{hancock2001rethinking}. The H index is an estimate of how well a set of items represents a latent variable. H values closer to 1 indicate a well-defined latent variable with factor loadings more likely to be stable across studies. In contrast, small H values suggest a poorly defined latent variable where factor loadings are liable to change across studies. Because we are primarily focused on measuring the reliability of MACE total and subscale/composite scores, we report the H values below but do not discuss them further.

@@

In addition to confirmatory factor analysis, we also we also performed exploratory factor analysis (EFA) on the response data from each sample separately. The goal of the EFA was to observe what factor structure would emerge with no \emph{a priori} restrictions \citep{schmitt2018selecting}. We decided to extract two- and three- factor solutions based on the total number of factors in the bifactor S-1 models. We were particularly interested in seeing if a two-factor solution would produce the factors from model 3 (i.e. threat and deprivation); and if a three-factor solution would produce the three factors from model 4 (i.e. maltreatment by adults, peer victimization, reverse-scoring).

@@

The goodness-of-fit indices for each model and sample are summarized in Table \ref{table:cfa_diagnostics}. According to the traditional indices, each model provided at least an acceptable fit to the response data. All RMSEA value were <0.05 and the majority of CFI and TLI values were >0.95. The only exception were for the fits of models 1 and 3 to the original sample data, which exhibited only acceptable fits (i.e. CFI/TLI > 0.90). In general, model fits were better for the replication sample data than for the original sample. Furthermore, all of the posterior predictive p-values corresponding to the $\chi^2_{NC}$ discrepancy measure were within the acceptable range. Thus all models were sufficiently able to reproduce the distribution of overall severity scores for each sample.

@@

In the following sections, we interpret the fit of each confirmatory item response model. To ease interpretation, we have reproduced the standardized factor loadings of each model for the original sample in Figure \ref{fig:loadings_original} and for the replication sample in Figure \ref{fig:loadings_online}. The model-based reliability indices for the bifactor models are summarized in Table \ref{table:reliability}.

@@

\subsubsection{Model 1: Unidimensional model of the MACE}

>>>>

% decrease size of references (lol)
\renewcommand*{\bibfont}{\small}

@@

\abstract{

}

@@

Childhood maltreatment is an unfortunately common experience for children across the globe \citep{stoltenborgh2015prevalence}. It is a risk factor for multiple poor outcomes later in life, including the onset of mental illness \citep{kessler2010childhood}, challenges in cognitive functioning \citep{su2019does}, and poorer physical health \citep{wegman2009meta}. Childhood maltreatment also predicts worse treatment outcomes for several psychiatric disorders \citep{nanni2012childhood, thomas2019childhood}. Thus, the reliable and valid assessment of childhood maltreatment is an important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design better interventions to prevent maltreatment in the first place.

There already exist many measures of childhood abuse and maltreatment for researchers to choose from \citep{saini2019systematic}. One recently developed measure, the Maltreatment and Abuse Chronology of Exposure scale (MACE; \citealt{teicher2015maltreatment}), has a number of notable advantages. The MACE is a 52-item retrospective self-report scale composed of 10 subscales, each of which measures a unique type of childhood maltreatment. Although six of these subscales overlap with other standard measures of abuse (e.g. physical abuse, neglect), the remaining four are less commonly represented in other scales (e.g. peer victimization, witnessing domestic violence). The items in the MACE were carefully selected to measure maltreatment of varying severity in order to finely differentiate people with differing exposures. The MACE also asks for the onset and duration of maltreatment experiences, which is necessary for investigating how the timing of abuse shapes development. The MACE was originally developed to provide two measures: a total score (number of overall maltreatment experiences) and a multiplicity score (number of types of maltreatment experienced). These scores have excellent temporal stability, convergent validity with other childhood maltreatment scales, and good predictive validity for a multitude of psychiatric symptoms \citep{teicher2015maltreatment}. For these reasons the MACE has been included as among the best measures of childhood maltreatment currently available \citep{saini2019systematic, georgieva2022systematic}.

In recent years, there have been calls for childhood maltreatment research to move away from the cumulative risk framework \citep{evans2013cumulative} --- which focuses only on the number of adverse experiences --- towards a dimensional approach, which emphasizes how distinct types of maltreatment uniquely shape developmental trajectories \citep{mclaughlin2016beyond, belsky2012beyond}. In response, some researchers have begun to use the MACE to form new indices of maltreatment in lieu of, or in addition to, the total and multiplicity scores. For example, some have used all 10 MACE subscale scores to study the unique contributions of each type of maltreatment to the development of psychopathology \citep{schalinski2015type, gerke2018childhood, schalinski2019early}. Inspired by a prominent dimensional model of maltreatment \citep{mclaughlin2014childhood}, others have made new subscale scores indexing threat and deprivation experiences using the items in the MACE that respectively measure abuse and neglect \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}. The psychometric properties of these new subscale scores have not been thoroughly investigated.

A critical challenge in studying the consequences of childhood maltreatment is the high co-occurrence of adverse experiences. That is, children who have experienced any single form of maltreatment are likely to have experienced multiple other types \citep{dong2004interrelatedness, herrenkohl2009assessing, kessler2010childhood}. As a consequence, isolating the unique contribution of one type of maltreatment to some outcome is difficult insofar as any one type of adversity is likely confounded with overall maltreatment. Concretely, this means that researchers intending to study the effects of particular types of maltreatment using subscale scores need to verify that those measures actually reflect a construct unique from from overall maltreatment, lest they draw spurious conclusions.

Bifactor models, a type of hierarchical factor model, are well-suited for addressing this issue \citep{bornovalova2020appropriate}. They assume that the covariation in responses to a set of items can be accounted for by a \emph{general} factor, reflecting shared variance across all items, and one or more \emph{specific} factors, reflecting additional shared variance among subsets of items \citep{Reise2012-ql}. Crucially, bifactor models can be used to calculate model-based reliability statistics, including the omega hierarchical ($\omega_h$) and omega subscale ($\omega_s$) indices \citep{reise2013scoring}, which respectively measure the proportion of variance in total and subscale scores attributable to the general factor and specific factors. Applied to the MACE, these indices can help determine whether subscale scores measuring particular types of maltreatment are interpretable as such, or if they instead chiefly reflect overall maltreatment.

@@

The participants in the current study (N=1633 total) belonged to one of two samples. The first sample of participants (N=1051) are from the original development of the MACE \citep{teicher2015maltreatment}. The full recruitment details for this sample have been reported elsewhere \citep{teicher2015maltreatment}. Briefly, these participants were recruited from the greater Boston area between 2010 and 2013. Inclusion criteria included being medically healthy, unmedicated, and between 18–25 years of age. This sample will hereafter be referred to as the \textit{original} sample.

@@

The demographics of the two samples are summarized in Table \ref{tab:demographics}. The majority of participants in both samples identified as women, though the original sample was composed of proportionally more woman than the replication sample ($z$ = 4.142, $p$ < 0.001). Participants in the original sample were also younger on average ($t$ = -19.313, $p$ < 0.001). The majority of participants in both samples identified as White and not as Hispanic or Latino. The proportion of participants identifying as White was not significantly different across samples ($z = 0.417$, $p = 0.677$), nor was the proportion of participants identifying as Hispanic or Latino ($z$ = -1.241, $p$ = 0.214).

@@

All participants completed the 52-item version of the MACE \citep{teicher2015maltreatment}. The scale is a retrospective self-report measure of the severity and timing of exposure to 10 types of childhood maltreatment. For each item, participants endorse whether they ever experienced a particular event during childhood (e.g. ``Parents or guardians intentionally pushed, grabbed, shoved, slapped, pinched, punched or kicked you'') and, if so, at what ages (up to age 18). Here we only analyze the dichotomous endorsement responses (Yes = 1, No = 0) and not the chronology data. Six of the 52-items --- evenly divided between the emotional and physical neglect subscales --- are positively valenced and reverse-scored (e.g. ``One or more individuals in your family helped you feel important or special''). For these items we follow the scoring convention set by \cite{teicher2015maltreatment}: a score of 1 is assigned to a response only in the absence of a positive event for all 18 years of childhood. This is in contrast to the remaining items in the scale, where a score of 1 is assigned to a response if maltreatment or abuse occurred in at least one year of childhood.

% (i.e. verbal abuse, physical abuse, nonverbal emotional abuse, sexual abuse, emotional neglect, and physical neglect by parents or guardians)

In both the original and replication samples, participants completed a number of additional self-report symptom measures. We do not include these in our analyses and therefore do not discuss these measures further in the main text. For completeness, we report these secondary self-report measures in the supplementary materials.

@@

To ensure data quality in the replication sample, we relied on attention checks embedded in the self-report measures to identify and remove participants engaging in careless or insufficient effort responding (\citealt{zorowitz2021inattentive}; see the supplementary materials for the attention checks). The data from N=64 (9.9\%)  participants in the replication sample were excluded for failing one or more of these attention checks, leaving the data from N=582 participants for analysis.

@@

Prior to analysis, we investigated whether it was appropriate to combine response data from the original and replication samples. We first compared maltreatment endorsement rates for each item between samples. The Spearman-rank correlation of endorsement rates across samples was large ($\rho$ = 0.951, p < 0.001), indicating excellent correspondence in the relative prevalence of maltreatment types across samples. When we compared the distribution of person-level total scores across samples, however, we found that scores were greater on average in the replication sample than in the original sample ($p_1$ = 0.185, $p_2$ = 0.219, $t$ = -4.611, p < 0.001).

An overall shift in overall endorsement rates by sample is not necessarily evidence for a violation of measurement invariance. One possibility is the replication sample had simply experienced more adversity than the original sample. A second is that the differences in the demographic composition of the two samples (e.g. gender, age) explain the differences in endorsement rates. To investigate these possibilities, we tested for uniform differential item functioning (DIF) using logistic regression, in which a participant's endorsement response on an item is regressed against their rest score (i.e. their total score for all other items), sample membership, gender, and age.

Using the cutoffs recommended by \cite{hidalgo2014binary}, we identified 11 of 52 items on the MACE that exhibited large DIF by study (Figure \ref{fig:dif_study}). Interestingly these included all six reverse-scored items. We also found an additional 11 items that exhibited large DIF by gender (Figure \ref{fig:dif_gender}). Five of these items measured sexual abuse (endorsed by more women than men), and another five measured peer physical abuse (endorsed by more men than women). We identified no items that exhibited large DIF by age. Given the number of items exhibiting DIF by sample membership, we decided to analyze each sample separately.

@@

The MACE contains a total of 12 item pairs and triplets characterized by response dependence (Table \ref{tab:dependence}). The items in some of these sets are structurally dependent, where a response to one item requires or strongly implies a certain response on the second. For example, a person endorsing item 8 (``Parents or guardians hit you so hard that you received medical attention'') is all but certain to endorse item 7 (``Parents or guardians hit you so hard that it left marks for more than a few minutes''). Other items are dependent by virtue of redundancy. One such example is items 36 (``Peers forced you to engage in sexual activity against your will'') and 37 (``Peers forced you to do things sexually that you did not want to do``); here the same question is asked twice using slightly different language.

We were concerned about response dependence for two reasons. First, it can inflate the estimates of item discrimination parameters, making items appear more informative than they actually are \citep{marais2008formalizing}. Second, response dependence can inflate the estimates of item loadings on specific factors in bifactor models, giving false support for models with a greater number of factors than needed \citep{reise2013applying}. Following \cite{marais2008formalizing}, we therefore converted each set of dependent items into a single polytomous item thereby reducing the MACE from 52 to 39 items in total.

@@

To investigate the factor structure of the MACE, we fit a series of confirmatory item factor models separately to the response data from each sample. The basis of each model is the graded response model for ordered polytomous data \citep{samejima1997graded}. The factor structure for each of the four models is depicted in Figure \ref{fig:models}. The first model is a one-factor (undimensional) model where all items load onto a general maltreatment factor (Figure \ref{fig:models}a). The second is a bifactor model where all items load onto a general maltreatment factor and also one of 10 specific factors based on their subscale membership (Figure \ref{fig:models}b).

The third model we fit investigated a conceptual model of maltreatment that distinguishes between threat and deprivation as distinct forms of childhood adversity \citep{mclaughlin2014childhood}. Initially we fit a bifactor model where all items loaded onto a general maltreatment factor and either the threat or deprivation specific factors. However, we observed vanishing item loadings on the threat specific factor. As such, we instead fit a bifactor S-1 model \citep{eid2017anomalous} where all items load onto a general maltreatment (or reference) factor, and only the items measuring neglect experiences load onto their own specific factor (Figure \ref{fig:models}c).

@@

The fourth and final model proposes a more parsimonious factor structure for the MACE. Items in the MACE can be neatly divided according to the source of adversity: maltreatment by adults at home (e.g. parents) or by peers. We hypothesized then that there would be two principal factors underlying responses on the MACE, maltreatment by adults and peer victimization. Based on preliminary model fits, we also suspected the presence of a secondary methods factor affecting the 6 reverse-scored items. Initially we fit a bifactor model in which all items loaded onto a general maltreatment factor and also one of three specific factors: maltreatment by adults, peer victimization, or reverse-scoring. However, we also observed vanishing loadings for the maltreatment by adults specific factor. Thus, we instead fit a bifactor S-1 model where all items load onto a general maltreatment factor; in addition, the peer victimization and reverse-scoring items load onto their own respective specific factors (Figure \ref{fig:models}d).

@@

We relied on multiple indices to judge the fit of the models to the data. First we calculated three traditional goodness-of-fit statistics using the ordinal $M_2$ statistic \citep{cai2013limited}. $M_2$ is asymptotically $\chi^2$-distributed with $\kappa − \nu$ degrees of freedom ($df$), where $\kappa$ is the reduced number of first- and second-order marginal residuals (i.e. $\kappa = n(n + 1)/2$ for $n$ items) and $\nu$ denotes the number of free parameters. To improve the stability of the $M_2$ statistic, the responses to polytomous items were binarized (i.e. none vs. any maltreatment). Using the $M_2$ statistic, we calculated the root mean square error of approximation (RMSEA), comparative fit index (CFI), and Tucker-Lewis index (TLI) for each model fit. Following convention \citep{hu1999cutoff}, the benchmarks for adequate model fit are RMSEA < 0.08, CFI > 0.90, and TLI > 0.90; in turn, the benchmarks for good model fit are RMSEA < 0.05, CFI > 0.95, and TLI > 0.95.

Given the limitations of SEM-based fit indices for item response models (e.g. \citealt{reise2014evaluating}), we calculated two additional fit indices. First, we used posterior predictive model checking to calculate a $\chi^2_{NC}$ discrepancy measure based on the total score distribution \citep{sinharay2006posterior}. This measure compares the observed and model-predicted proportion of participants at each total score level. Second, we calculated Yen's $Q_3$ statistic for each pair of items to detect local dependence \citep{yen1984effects} and also the critical $Q_3$-value above which local dependence is indicated  \citep{christensen2017critical}. Briefly, for each model and sample, we simulated 1000 locally independent datasets using the posterior distribution of the model parameters. We then recorded the maximum $Q_3$ value per simulation. The critical $Q_3$-value was defined as the 99th percentile of these max $Q_3$ values.

@@

To evaluate the reliability of the MACE total and subscale scores, we computed several model-based reliability indices using the factor loading estimates from the bifactor models. First, we calculated coefficient omega ($\omega$; \citealt{mcdonald1999test}), which quantifies the proportion of reliable score variance attributable to all factors (i.e. both group and specific factors). Next we calculated coefficient omega hierarchical ($\omega_h$) and coefficient omega subscale ($\omega_s$;  \citealt{reise2013scoring}). Whereas $\omega_h$ measures the fraction of total score variance attributable to the general factor, $\omega_s$ measures the fraction of subscale score variance attributable to a specific factor after controlling for the general factor. The values of $\omega_h$/$\omega_s$ facilitate interpretation of total and subscale scores. Values of $\omega_h$ approaching 1 indicate the general factor is the primary source of reliable variance in a total score. When $\omega_h$ > 0.80, the total score may be considered an essentially unidimensional reflection of the general factor even if the data are multidimensional \citep{rodriguez2016applying}. So too values of $\omega_s$ approaching 1 indicate the specific factor, and not the general factor, is the primary source of reliable variance in a subscale score. \cite{canivez2016bifactor} suggested that an acceptable $\omega_s$ value is 0.50 and that >0.70 is desirable. In contrast, values of $\omega_s$ smaller than 0.50 preclude the interpretation of subscale scores as reflecting a specific factor \citep{gignac2013bifactor}. 
To assess the unidimensionality of the MACE, we also calculated the explained common variance (ECV; \citealt{sijtsma2009use}) and the proportion of uncontaminated correlations (PUC; \citealt{reise2013multidimensionality}). The ECV is the ratio of the variance explained by the general factor divided by the total common variance of a scale; it therefore measures the contribution of the general factor relative to the specific factors. As ECV values approach 1, the general factor loadings of a bifactor model are expected to resemble those that would be obtained by a one-dimensional model. ECV values greater than 0.70 often indicate a scale is essentially unidimensional \citep{rodriguez2016applying}. In turn, the PUC quantifies how many inter-item correlations are accounted for only by the general factor and is an important moderator of the ECV. As the PUC increases, ECV is less important for evaluating the risk of bias when fitting a unidimensional model to data with a latent bifactor structure. When PUC > 0.80, there is a low risk of bias when a multidimensional scale is treated as unidimensional \citep{reise2013multidimensionality}. As an additional check, we calculated the relative parameter bias (RPB), which is the difference between an item's loading in a unidimensional model and its corresponding general factor loading the bifactor, divided by the general factor loading in the bifactor. Parameter bias less than 10–15\% is acceptable \citep{muthen1987structural}.

Finally, we calculated the H index to assess the construct replicability of each factor \citep{hancock2001rethinking}. The H index is an estimate of how well a set of items represents a latent variable. H values closer to 1 indicate a well-defined latent variable with factor loadings more likely to be stable across studies. In contrast, small H values suggest a poorly defined latent variable where factor loadings are liable to change across studies. Because we are primarily focused on measuring the reliability of MACE total and subscale scores, we report the H values below but do not discuss them further.

@@

In addition to confirmatory factor analysis, we also we also performed exploratory factor analysis (EFA) on the response data from each sample separately. The goal of the EFA was to observe what factor structure would emerge with no \emph{a priori} restrictions. We decided to extract two- and three- factor solutions based on the total number of factors in the bifactor S-1 models. We were particularly interested in seeing if a two-factor solution would produce the factors from model 3 (i.e. threat and deprivation); and if a three-factor solution would produce the three factors from model 4 (i.e. maltreatment by adults, peer victimization, reverse-scoring).

@@

The goodness-of-fit indices for each model and sample are summarized in Table \ref{table:cfa_diagnostics}. According to the traditional indices, each model provided at least an acceptable fit to the data. All RMSEA value were <0.05 and most CFI and TLI values were >0.95. The only exception were for the fits of models 1 and 3 to the original sample data where CFI/TLI > 0.90. In general, model fits were better for the replication sample data than for the original sample. Furthermore, none of the posterior predictive p-values corresponding to the $\chi^2_{NC}$ discrepancy measure exceeded the critical values indicating that all models were able to reproduce the observed distribution of total scores for each sample.

@@

In the following sections, we interpret the fit of each confirmatory item response model. To facilitate interpretation, we have reproduced the standardized factor loadings of each model for the original sample in Figure \ref{fig:loadings_original} and for the replication sample in Figure \ref{fig:loadings_online}. The model-based reliability indices for the bifactor models are summarized in Table \ref{table:reliability}.

@@

\subsubsection{Model 1: One-factor (unidimensional) model of the MACE}
